On The Statistical Properties of Cospectra

### 2017-12-18 paper/cospectra-paper.tex
New section on white noise simulations with dead time

Time series analysis is one of the primary ways to understand the physical properties of astronomical object in our universe, from exoplanets and stars to black holes and Active Galactic Nuclei (AGN). 
Fourier analysis, especially through the periodogram\footnote{We distinguish in this paper between the \textit{power spectrum}, which describes the process at the source generating variable time series, and the \textit{periodogram}, which denotes a realization of said power spectrum, i.e.\ the time series we actually observe. \textbf{In line with the signal processing literature, we also use the term \textit{periodogram} or \textit{power spectral density} for the square of the absolute value of Fourier amplitudes derived from an observed light curve}}, has long been used to find periodic and quasi-periodic signals as well as characterize the stationary stochastic processes often present in accreting systems. 
While in principle, the statistics of the periodogram is well understood and characterized in the literature \citep[e.g.][]{vanderklis1989}, the periodogram is often subject to instrumental effects like dead time that change its statistical properties and thus make statistical inference difficult in practice \citep[e.g.][]{Zhang+95}.

@@

Within astronomy, it has recently been recognized as one approach to mitigate instrumental effects by making use of the independent detectors present in some telescopes. Where in standard analysis, light curves of multiple detectors are summed before Fourier transforming the summed light curve, it is possible to instead Fourier-transform the signal of two independent detectors \textbf{within the same instrument observing the same source strictly simultaneously in the same energy band}. 
The resulting \textbf{co-spectral} densities will be less affected by an instrumental effect called \textit{dead time} (see details in \citealt{Bachetti+15}). \textbf{In many X-ray detectors, after detection of a photon there is a time interval during which the detector (or in imaging detectors the individual pixel) is blocked from detecting a second photon. This time interval is generally called \textit{dead time} (the detector is effectively ``dead'', and incoming photons will not produce a signal) and its characteristic time scale is set by the details of the detector and can range from very short ($2\mu\mathrm{s}$ in the \fermi\ Gamma-Ray Burst Monitor) to very long ($2.5\mathrm{ms}$ in \nustar).} It leads to frequency-dependent changes in the mean and variance of the statistical distributions governing the periodogram, \textbf{which cannot be mitigated by averaging periodograms of multiple segments}. This is relevant for current X-ray missions \textbf{carrying at least two identical detectors} such as \nustar\ and \astrosat, but will also be relevant to future missions with multiple detectors like \ixpe. This approach has recently been used in \nustar\ studies of millisecond pulsars \citep{ferrigno2017}, Ultraluminous X-ray Sources \citep{bachetti2016} and X-ray binaries \citep{barriere2015,zoghbi2016,ingram2016,huppenkothen2017,stiele2017}. \textbf{It is particularly important for instruments like \nustar\ which have a fairly long dead time and thus exhibit dead time effects in the periodogram at frequencies comparable to those where signals are expected, and at fairly low count rates}.  
Similarly, the cospectrum of two time series taken with the same instrument, but in different wavelength ranges, can be used to characterize the coherent, \textbf{phase-aligned} variability in both time series as a function of frequency.

@@

In this paper, we lay out the basic statistical distributions for detecting periodic and narrow quasi-periodic signals in the presence of detector noise (e.g.\ photon counting statistics) for both single and averaged cospectra. We show below that unlike for the periodogram, the statistical distribution for a single cospectrum reduces to a Laplace distribution, while the distribution for averaged cospectra is considerably more complex. We also find that for averaged cospectra consisting of more than $\sim$30 averaged individual segments, the assumption of a Gaussian distribution is indeed appropriate for single-trial tail probabilities and reduces computation overhead. However, for averaged cospectra of fewer segments, more stringent significance thresholds or large numbers of trials, the statistical distribution--and hence the derived $p$-values used for period detection--deviate significantly from a Gaussian distribution. \textbf{We note that the results below hold only for cases where the signal to be detected is phase-aligned in both light curves. This must be true by definition for analyses utilizing data from independent, identical detectors observing the same source simultaneously, but might not strictly be true for observations of the same source e.g.\ in different energy bands. We caution the reader that for signals that are not phase-aligned, the resulting power may decrease significantly and result in a non-detectable signal.}

@@

\textbf{In the following, we will consider the example of detecting a strictly periodic or very narrow quasi-periodic signal in the presence of simple white noise, as is commonly the case in pulsar searches in X-ray data. We assume that both the white noise and the periodic signal are wide-sense stationary and the light curve is evenly sampled. While the white noise is also assumed to be strictly stochastic, the signal to be detected may be either deterministic (in the case of a strictly periodic signal) or stochastic (for a quasi-periodic oscillation). We then aim to reject a null hypothesis where the observed cospectral density at a given frequency can be explained by white noise alone.}

@@

\noindent For \textbf{evenly sampled}, normally distributed light curves (and indeed for uncertainties coming from a wide range of statistical distributions, including the Poisson distribution, if $N$ is large), the real and imaginary amplitude components are distributed as $A_{xj}, B_{xj} \sim \mathcal{N}(0, \sigma_x^2)$ with $\sigma_x =  \sqrt{\frac{\sum_{k=1}^{N}{x_k}}{2}}$ and  $A_{yj}, B_{yj} \sim \mathcal{N}(0, \sigma_y^2)$ with $\sigma_y = \sqrt{\frac{\sum_{k=1}^{N}{y_k}}{2}}$. \textbf{Note that these distributions for the Fourier amplitudes only hold if the underlying process producing the observations is stochastic and wide-sense stationary (like e.g.\ the white noise case considered here) as is commonly observed in a wide range of astrophysical sources including black hole X-ray binaries and AGN. For strictly deterministic processes, the distributions of the Fourier amplitudes will not be centered on $\mu = 0$, and the distributions below will not be correct.}

@@

\begin{equation}
M_Z(t) = \frac{\exp{\left( \frac{t\mu_x \mu_y + 0.5 (\mu_y^2 \sigma_x^2 + \mu_x^2 \sigma_y^2) t}{1 - t^2 \sigma_x^2 \sigma_y^2} \right)}}{\sqrt{1 - t^2 \sigma_x^2 \sigma_y^2}}\, ,
\end{equation}
but since $\mu_x = \mu_y = 0$ for Fourier amplitudes, this reduces to
\begin{equation}
M_Z(t) =  \frac{1}{\sqrt{1 - t^2 \sigma_x^2 \sigma_y^2}}  \label{eqn:mgf},
\end{equation}

@@

Detection thresholds for cospectra will generally be different from those of classical periodograms, because the Laplace distribution tends to be narrower than the equivalent $\chi^2_2$ distribution for single periodograms. To show how the distributions and the corresponding detection thresholds differ, we simulated simple Poisson-distributed light curves. First, we simulated two light curves with a duration of $10\,\mathrm{s}$ and $10^6$ data points each, corresponding to a time resolution of $10^{-5}\,\mathrm{s}$. The light curves have an identical mean count rate of $10^{6} \, \mathrm{counts/s}$, corresponding to $10$ counts per bin. In order to simulate typical measurement uncertainties, we sampled from a Poisson distribution for each time bin with a rate parameter $\lambda = 10$, corresponding to the average counts per bin.

@@

@@

\textbf{All caveats applying more generally to Fourier analysis and to periodograms specifically also apply to cospectra. In particular, as for the $\chi^2$ distribution used for period searches in periodograms, the distributions derived here assume that the underlying process is wide-sense stationary and the light curves are simultaneously and evenly sampled. These conditions are typically met in most observations taken with X-ray timing instruments like \nustar\ and \rxte\ for problems such as timing of X-ray binaries, AGN and X-ray pulsar searches. Note that this assumes, however, that the time scales of interest are much shorter than outburst timescales in X-ray binaries or changes in the X-ray background in pulsar searchers.}

\textbf{Non-stationarity (as is e.g.\ observed in flares) will impose a window function onto the cospectrum and shift the mean of the Fourier amplitudes, thus invalidating the assumptions above (for an illustrative example of this effect and its implications on period searches in periodograms of magnetar bursts, see \citealt{huppenkothen2013}). Uneven sampling more generally introduces covariances between adjacent frequencies and renders many of the statistical assumptions underpinning period searches using both in the periodogram and the cospectrum invalid. For unevenly sampled data, other methods that do not depend on a regular sampling pattern must be employed (for example the Lomb-Scargle periodogram \citep{lomb1976, scargle1982} or the Bayesian period search in \citet{gregory1992}).}

@@

\section{Example: A (Q)PO in \nustar}
\label{sec:nustarqpo}

In the following, we first simulate a QPO as expected to be observed in a typical \nustar\ observation, and then show similarly real \nustar\ data containing a coherent pulsed signal.

\subsection{Simulated Data}
In order to show the difference of the detection limits with the cospectrum and the power spectrum, we show how a QPO at 200 Hz from a very bright source would appear in \nustar. 
We simulate a light curve of $T=3000\mathrm{s}$ duration with a time resolution of $\delta t = 0.5 \,\mathrm{ms}$ and an average count rate of $200 \,\mathrm{counts}\,\mathrm{s}^{-1}$. To this constant background we add a quasi-periodic oscillation with a period of $5\,\mathrm{ms}$, a fractional rms amplitude of $f_\mathrm{rms} = 0.15$ and phases randomized using a normal distribution with a mean of $0$ and a width of $\sigma_\mathrm{qpo} = 0.01$. After producing this light curve, the procedure is similar to that followed by \citet{Bachetti+15}. We simulated photon events using rejection sampling from this light curve using the software package \textit{stingray}\footnote{\url{https://github.com/StingraySoftware/stingray}}, running the function \texttt{stingray.Eventlist.simulate\_times()} twice in order to produce two light curves that are statistically independent, but have the same signal and properties, as we would expect from an instrument with two independent detectors observing the same object. Subsequently we simulated variable deadtime for both light curves with an average time scale of $2.5 \,\mathrm{ms}$ as commonly seen in \nustar\ data \citep{Bachetti+15} using \textit{HENDRICS}\footnote{\url{https://github.com/StingraySoftware/HENDRICS}} \citep{bachetti2015b}. We then produced the periodogram of the summed light curves, the averaged periodogram of the two individual light curves, and the cospectrum of the two light curves. Note that in all three cases we produced averaged periodograms and co-spectra by splitting the light curves into 600 segments of $5\,\mathrm{s}$ length each in order to suppress the variance in the powers and show the effects of dead time more clearly.

@@

\subsection{\nustar\ Observations}

\begin{bf}
Hercules X-1 (Her X-1) is a well-studied persistent X-ray binary pulsars \citep{tananbaum1972} in an X-ray binary comprising the neutron star itself and a stellar companion HZ Herculis \citep{davidsen1972,forman1972,bahcall1972} with a mass of $\sim$$2.2 \,\mathrm{M}_\odot$ \citep{reynolds1997,leahy2014} whose type varies between late-type A and early-type B with orbital phase \citep{anderson1994,cheng1995} . The neutron star spins with a period of $P = 1.23\,\mathrm{s}$ \citep{tananbaum1972} and the system overall exhibits an orbital period of $P_\mathrm{orb} = 1.7\,\mathrm{days}$ \citep{bahcall1972}, along with super-orbital variations on a $\sim$35-day timescale \citep{giacconi1973,scott1999,igna2011}.
q
Her X-1 was observed by the Nuclear Spectroscopic Telescope Array (\nustar) multiple times in the last years. 
For this work, we considered the observation taken from UT 2012-09-19 to UT 2012-09-20, one of those used by \citet{fuerst13} to characterize the cyclotron resonance scattering features in the spectrum of the source.
We downloaded the observation directory for observation ID 30002006002 from the HEASARC and used the FTOOL \texttt{barycorr} on the L2 cleaned science event files (file name ending with \verb|01_cl.evt|) to correct the photon arrival times to the solar system barycenter.
For our analysis, we considered photons from 3 to 79 keV at most 50\arcsec\ from the nominal position of the source, extracted from the two identical \textit{Focal Plane Modules A} and \textit{B} (FPMA and FPMB, respectively) onboard the spacecraft.
For this work, we used a total of $32.67\,\mathrm{ks}$ of good time intervals (GTIs), only selecting intervals longer than $10\,\mathrm{s}$. In Figure \ref{fig:herx1_lc} (upper panel), we present the light curve of the \nustar Her X-1 observations. The source varied substantially in brightness during the observations on fairly long time scales (10s to 1000s of seconds), indicating a significant source of red noise at low frequencies.

>>>>

Time series analysis is one of the primary ways to understand the physical properties of astronomical object in our universe, from exoplanets and stars to black holes and Active Galactic Nuclei (AGN). 
Fourier analysis, especially through the periodogram\footnote{We distinguish in this paper between the \textit{power spectrum}, which describes the process at the source generating variable time series, and the \textit{periodogram}, which denotes a realization of said power spectrum, i.e.\ the time series we actually observe. \textbf{In line with the signal processing literature, we also use the term \textit{periodogram} or \textit{power spectral density} for the square of the absolute value of Fourier amplitudes derived from an observed light curve. We use the terms \textit{cospectrum} or \textit{cospectral density} to denote the real component of the cross-spectrum, i.e.\ the result of a multiplication of the Fourier amplitudes of one light curve with the complex conjugate of the Fourier amplitudes of a second light curve (see also Section \ref{sec:whitenoise_cospectra} for an exact definition)}}, has long been used to find periodic and quasi-periodic signals as well as characterize the stationary stochastic processes often present in accreting systems. 
While in principle, the statistics of the periodogram is well understood and characterized in the literature \citep[e.g.][]{vanderklis1989}, the periodogram is often subject to instrumental effects like dead time that change its statistical properties and thus make statistical inference difficult in practice \citep[e.g.][]{Zhang+95}.

@@

Within astronomy, it has recently been recognized as one approach to mitigate instrumental effects by making use of the independent detectors present in some telescopes, in particular an effect called \textit{dead time}.  \textbf{In many X-ray detectors, after detection of a photon there is a time interval during which the detector (or in imaging detectors the individual pixel) is blocked from detecting a second photon. This time interval is generally called \textit{dead time} (the detector is effectively ``dead'', and incoming photons will not produce a signal) and its characteristic time scale is set by the details of the detector and can range from very short ($2\mu\mathrm{s}$ in the \fermi\ Gamma-Ray Burst Monitor) to very long ($2.5\mathrm{ms}$ in \nustar).} It leads to frequency-dependent changes in the mean and variance of the statistical distributions governing the periodogram, \textbf{which cannot be mitigated by averaging periodograms of multiple segments}.

Where in standard analysis, light curves of multiple detectors are summed before Fourier transforming the summed light curve, it is possible to instead Fourier-transform the signal of two independent detectors \textbf{within the same instrument observing the same source strictly simultaneously in the same energy band}. 
The resulting \textbf{co-spectral} densities will be less affected by dead time (see details in \citealt{Bachetti+15}), because the latter is introduced in each detector independently and tends to cancel out in the cospectrum.
This approach relevant for current X-ray missions \textbf{carrying at least two identical detectors} such as \nustar\ and \astrosat, but will also be relevant to future missions with multiple detectors like \ixpe. It has recently been used in \nustar\ studies of millisecond pulsars \citep{ferrigno2017}, Ultraluminous X-ray Sources \citep{bachetti2016} and X-ray binaries \citep{barriere2015,zoghbi2016,ingram2016,huppenkothen2017,stiele2017}. \textbf{It is particularly important for instruments like \nustar\ which have a fairly long dead time and thus exhibit dead time effects in the periodogram at frequencies comparable to those where signals are expected, and at fairly low count rates}.  
Similarly, the cospectrum of two time series taken with the same instrument, but in different wavelength ranges, can be used to characterize the coherent, \textbf{phase-aligned} variability in both time series as a function of frequency.

@@

In this paper, we lay out the basic statistical distributions for detecting periodic and narrow quasi-periodic signals in the presence of detector noise (e.g.\ photon counting statistics) for both single and averaged cospectra. We show below that unlike for the periodogram, the statistical distribution for a single cospectrum reduces to a Laplace distribution, while the distribution for averaged cospectra is considerably more complex. We also find that for averaged cospectra consisting of more than $\sim$30 averaged individual segments, the assumption of a Gaussian distribution is indeed appropriate for single-trial tail probabilities and reduces computation overhead. However, for averaged cospectra of fewer segments, more stringent significance thresholds or large numbers of trials, the statistical distribution--and hence the derived $p$-values used for period detection--deviate significantly from a Gaussian distribution. \textbf{We note that the results below hold only for cases where the signal to be detected is phase-aligned in both light curves. This must be true by definition for analyses utilizing data from independent, identical detectors observing the same source simultaneously (unless detectors are so far apart that light travel time effects might be an issue and the data is not barycentred to the gravitational centre of the solar system), but might not strictly be true for observations of the same source e.g.\ in different energy bands. We caution the reader that for signals that are not phase-aligned, the resulting power may decrease significantly and result in a non-detectable signal.}

@@

\textbf{In the following, we will consider the example of detecting a strictly periodic or very narrow quasi-periodic signal in the presence of simple white noise, as is commonly the case for example in pulsar searches in X-ray data. We assume that both the white noise and the periodic signal are wide-sense stationary (i.e. the mean and the autocovariance of the time series do not change with time) and the light curve is evenly sampled simultaneously in two identical detectors. While the white noise is also assumed to be strictly stochastic, the signal to be detected may be either deterministic (in the case of a strictly periodic signal) or stochastic (for a quasi-periodic oscillation). We then aim to reject a null hypothesis where the observed cospectral density at a given frequency can be explained by white noise alone.}

@@

\noindent For \textbf{evenly sampled}, normally distributed light curves (and indeed for uncertainties coming from a wide range of statistical distributions, including the Poisson distribution, if $N$ is large), the real and imaginary amplitude components are distributed as $A_{xj}, B_{xj} \sim \mathcal{N}(0, \sigma_x^2)$ with $\sigma_x =  \sqrt{\frac{\sum_{k=1}^{N}{x_k}}{2}}$ and  $A_{yj}, B_{yj} \sim \mathcal{N}(0, \sigma_y^2)$ with $\sigma_y = \sqrt{\frac{\sum_{k=1}^{N}{y_k}}{2}}$. \textbf{Note that these distributions for the Fourier amplitudes only hold if the underlying process producing the observations is stochastic and wide-sense stationary. This includes many processes commonly observed in astrophysical sources such as white noise observed from a constant background, as well as red noise processes such as shot noise and other (broken) power-law power spectra often seen in AGN, and quasi-periodic oscillations with stochastic variations in amplitude and period of the observed signal, commonly observed in black hole X-ray binaries. For strictly deterministic processes (e.g.\ strictly periodic variations), the distributions of the Fourier amplitudes will not be centered on $\mu = 0$, and the distributions below will not be correct.}

@@

\begin{equation}
M_Z(t) = \frac{\exp{\left( \frac{t\mu_x \mu_y + 0.5 (\mu_y^2 \sigma_x^2 + \mu_x^2 \sigma_y^2) t}{1 - t^2 \sigma_x^2 \sigma_y^2} \right)}}{\sqrt{1 - t^2 \sigma_x^2 \sigma_y^2}}\, ,
\end{equation}
but since $\mu_x = \mu_y = 0$ for the Fourier amplitudes \textbf{of stationary stochastic processes}, this reduces to
\begin{equation}
M_Z(t) =  \frac{1}{\sqrt{1 - t^2 \sigma_x^2 \sigma_y^2}}  \label{eqn:mgf},
\end{equation}

@@

Detection thresholds for cospectra will generally be different from those of classical periodograms, because the Laplace distribution tends to be narrower than the equivalent $\chi^2_2$ distribution for single periodograms. To show how the distributions and the corresponding detection thresholds differ, we simulated simple Poisson-distributed light curves. First, we simulated two light curves with a duration of $10\,\mathrm{s}$ and $10^6$ data points each, corresponding to a time resolution of $10^{-5}\,\mathrm{s}$. The light curves have an identical mean count rate of $10^{6} \, \mathrm{counts/s}$, corresponding to $10$ counts per bin. In order to simulate typical measurement uncertainties in X-ray detectors, we sampled from a Poisson distribution for each time bin with a rate parameter $\lambda = 10$, corresponding to the average counts per bin.

@@

\textbf{All caveats applying more generally to Fourier analysis and to periodograms specifically also apply to cospectra. In particular, as for the $\chi^2$ distribution used for period searches in periodograms, the distributions derived here assume that the underlying process is wide-sense stationary and the light curves are simultaneously and evenly sampled. These conditions are typically met in most observations taken with X-ray timing instruments like \nustar\ and \rxte\ for problems such as timing of X-ray binaries, AGN and X-ray pulsar searches, but this may not be the case for e.g.\ optical observations of main-sequence or binary stars. Note that we also implicitly assume that the time scales of interest are much shorter than outburst timescales in X-ray binaries or changes in the X-ray background in pulsar searchers, both of which may not be stationary.}

\textbf{Non-stationarity (as is e.g.\ observed in flares) will impose a window function onto the cospectrum and shift the mean of the Fourier amplitudes, thus invalidating the assumptions above (for an illustrative example of this effect and its implications on period searches in periodograms of magnetar bursts, see \citealt{huppenkothen2013}). Uneven sampling, on the other hand, introduces covariances between adjacent frequencies and renders many of the statistical assumptions underpinning period searches using both in the periodogram and the cospectrum invalid. For unevenly sampled data, other methods that do not depend on a regular sampling pattern must be employed (for example the Lomb-Scargle periodogram \citep{lomb1976, scargle1982} or the Bayesian period search in \citet{gregory1992}) and dead-time effects must be forward-simulated.}

\begin{figure*}
\begin{center}
\includegraphics[width=\textwidth]{whitenoise_deadtime.png}
\caption{\textbf{Left panel: Fourier products derived from white noise light curves heavily affected by dead time. There is a strong, frequency-dependent modulation in the periodogram (orange) introduced by dead time effects, rendering standard tests for periodicities unreliable. The cospectrum (dark grey) is less affected overall, but there remains a frequency-dependent variation in the local variance, with a narrower distribution governing the cospectral densities at low frequencies, and a wider distribution e.g.\ between 200 and 400 Hz. Applying the FAD correction (blue) produces a cospectrum that follows the distributions defined in this paper very closely, allowing for reliable periodicity testing. Right panel: histograms of cospectral densities derived from $1000$ simulated, dead time-affected cospectra (dark grey) and their FAD-corrected equivalents (blue). The dead time-affected cospectra overall have a narrower distribution (but note that this is integrated over all frequencies, thus the frequency dependence is washed out), whereas the corrected cospectra follow the expected distribution (red).
}}
\label{fig:wn_dt}
\end{center}
\end{figure*}

@@

\section{Example: \nustar}
\label{sec:nustarqpo}

\textbf{In the following, we will consider three more realistic examples in turn. First, we will consider simulations of white noise light curves with strong dead time effects. We will then provide more realistic simulations of a QPO as expected to be observed in a typical \nustar\ observation, and finally present an example of real \nustar\ data containing a coherent pulsed signal from the X-ray neutron star source Hercules X-1.}

\subsection{Simulated Light Curves With Dead Time}

\begin{bf}
Here, we simulate simple white noise light curves and apply a typical \nustar-like dead time of $2.5\mathrm{ms}$ in order to explore how the distributions defined above work under more realistic conditions. We use the relatively extreme case of the Crab pulsar, one of the brightest variable X-ray sources in the sky [REF]. We simulate event files for an observation of $T = 100\,\mathrm{s}$ and a constant count rate of $c = 400 \mathrm{counts}/\mathrm{s}$. Incident photon events are generated by picking $N=cT = 4 \times 10^{4}$ times of arrivals randomly from a uniform distribution between the start and end time, independently for each of the two \nustar\ detectors. We then apply dead time to both sets of photon events by excluding all photons that arrive within $2.5\mathrm{ms}$ of the previous photon arrival in each detector. For a count rate of $400\mathrm{counts}/\mathrm{s}$ and a \nustar-like dead time, this corresponds to a photon loss of $\sim$50\%. The resulting light curves are binned at a time resolution of $9.766 \times 10^{-4}\,\mathrm{s}$ (corresponding to a Nyquist frequency of $\nu_{\mathrm{Nyquist}} = 512 \,\mathrm{Hz}$) and averaged periodograms of each individual light curve, of the summed light curves, and the averaged cospectrum are created using segments of $t_\mathrm{seg} = 10\,\mathrm{s}$. We repeated this process 1000 times in order to gain sufficient simulated cospectral densities for statistical comparisons.

In Figure \ref{fig:wn_dt} (left panel), we show the periodogram, cospectrum and the corresponding probability distribution of cospectral densities of these simulations. In the periodogram, we see a large frequency-dependent deviation from the expected white noise level of $2$, which is difficult to correct for in practice and introduces a large frequency-dependent bias into any standard tests for periodicities that utilize the periodogram. The cospectrum, on the other hand, does not suffer from distortions of the underlying power spectrum, but does show frequency-dependent variations in the local variance (see also \citet{Bachetti+15} for a more in-depth explanation of this effect). 
This indicates that the distributions defined in this paper cannot be applied to dead-time affected data without correction. Here, we follow the suggestion of \citet{bachetti2017} and use so-called \textit{Fourier Amplitude Differences} (FAD) in order to correct the cospectrum. In short, this technique rests on the idea that while variability intrinsic to the source will produce Fourier amplitudes with the same real and imaginary components, both dead time and Photon counting noise are independent between detectors, and will thus be uncorrelated between Fourier amplitudes derived from two identical, independent detectors. The \textit{difference} between Fourier amplitudes of one detector and another will hence effectively remove source-intrinsic variability and only contain dead-time related effects. One can then model the latter in the Fourier amplitude differences, and use this model to correct the original cospectrum (see \citet{bachetti2017} for details).

We corrected the cospectra using the FAD method and subsequently compared the empirical distributions derived from the simulated cospectra to the theoretical distribution defined in Section \ref{sec:averaged_detthres} (Figure \ref{fig:wn_dt}, right panel). We find that the FAD-corrected cospectra follow the derived distributions very closely even in this case involving an exceptionally high degree of dead time, and conclude that for most realistic scenarios, the distributions arrived here can be used to construct unbiased estimators for periodicities, assuming that the FAD correction is applied to the cospectra first.

\end{bf}

@@

\subsection{A Simulated QPO in \nustar}

In order to show the difference of the detection limits with the cospectrum and the power spectrum, we show how a QPO at 200 Hz from a very bright source would appear in \nustar. 
We simulate a light curve of $T=3000\mathrm{s}$ duration with a time resolution of $\delta t = 0.5 \,\mathrm{ms}$ and an average count rate of $200 \,\mathrm{counts}\,\mathrm{s}^{-1}$. To this constant background we add a quasi-periodic oscillation with a period of $5\,\mathrm{ms}$, a fractional rms amplitude of $f_\mathrm{rms} = 0.15$ and phases randomized using a normal distribution with a mean of $0$ and a width of $\sigma_\mathrm{qpo} = 0.01$. After producing this light curve, the procedure is similar to that followed by \citet{Bachetti+15}. We simulated photon events using rejection sampling from this light curve using the software package \textit{stingray}\footnote{\url{https://github.com/StingraySoftware/stingray}}, running the function \texttt{stingray.Eventlist.simulate\_times()} twice in order to produce two light curves that are statistically independent, but have the same signal and properties, as we would expect from an instrument with two independent detectors observing the same object. Subsequently we simulated variable deadtime for both light curves with an average time scale of $2.5 \,\mathrm{ms}$ as commonly seen in \nustar\ data \citep{Bachetti+15} using \textit{HENDRICS}\footnote{\url{https://github.com/StingraySoftware/HENDRICS}} \citep{bachetti2015b}. We then produced the periodogram of the summed light curves, the averaged periodogram of the two individual light curves, and the cospectrum of the two light curves. Note that in all three cases we produced averaged periodograms and co-spectra by splitting the light curves into 600 segments of $5\,\mathrm{s}$ length each in order to suppress the variance in the powers and show the effects of dead time more clearly. \textbf{The cospectra were FAD-corrected after averaging.}

@@

\subsection{\nustar\ Observations of Her X-1}

\begin{bf}
Hercules X-1 (Her X-1) is a well-studied persistent X-ray binary pulsars \citep{tananbaum1972} in an X-ray binary comprising the neutron star itself and a stellar companion HZ Herculis \citep{davidsen1972,forman1972,bahcall1972} with a mass of $\sim$$2.2 \,\mathrm{M}_\odot$ \citep{reynolds1997,leahy2014} whose type varies between late-type A and early-type B with orbital phase \citep{anderson1994,cheng1995} . The neutron star spins with a period of $P = 1.23\,\mathrm{s}$ \citep{tananbaum1972} and the system overall exhibits an orbital period of $P_\mathrm{orb} = 1.7\,\mathrm{days}$ \citep{bahcall1972}, along with super-orbital variations on a $\sim$35-day timescale \citep{giacconi1973,scott1999,igna2011}.

Her X-1 was observed by the Nuclear Spectroscopic Telescope Array (\nustar) multiple times in the last years. 
For this work, we considered the observation taken from UT 2012-09-19 to UT 2012-09-20, one of those used by \citet{fuerst13} to characterize the cyclotron resonance scattering features in the spectrum of the source.
We downloaded the observation directory for observation ID 30002006002 from the HEASARC and used the FTOOL \texttt{barycorr} on the L2 cleaned science event files (file name ending with \verb|01_cl.evt|) to correct the photon arrival times to the solar system barycenter.
For our analysis, we considered photons from 3 to 79 keV at most 50\arcsec\ from the nominal position of the source, extracted from the two identical \textit{Focal Plane Modules A} and \textit{B} (FPMA and FPMB, respectively) onboard the spacecraft.
For this work, we used a total of $32.67\,\mathrm{ks}$ of good time intervals (GTIs), only selecting intervals longer than $10\,\mathrm{s}$. In Figure \ref{fig:herx1_lc} (upper panel), we present the light curve of the \nustar Her X-1 observations. The source varied substantially in brightness during the observations on fairly long time scales (10s to 1000s of seconds), indicating a significant source of red noise at low frequencies.
